### 2011-04-03 acronym.tex
second commit

\section*{Acronyms}
\begin{acronym}
%\acronym{JDT}{\emph{Java Development Tools}}
%\acronym{NFC}{\emph{Number of Function Calls}}
\setlength{\itemsep}{-\parsep} 
\acro{SP}[SP]{Standardized Patient}
\acro{SPs}[SPs]{Standardized Patients}
\acro{etext}[e-text]{electronic text}
\acro{pdf}[PDF]{Portable Document File}
\acro{CSVs}[CSVs]{Comma Separated Values }
\acro{ttcn3}[TTCN-3]{Test and Test Control Notation Version 3}
\acro{ttcn}[TTCN]{Tree and Tabular Combined Notation}
\acro{etsi}[ETSI]{European Telecommunications Standards Institute}
\acro{itut}[ITU-T]{International Telecommunication Union}
\acro{3gpp}[3GPP]{Third Generation Partnership Project}
\acro{ims}[IMS]{Ip Multimedia Systems}
\acro{itut}[ITU-T]{International Telecommunication Union}
\acro{iso}[ISO]{International Organization for Standardization}
\acro{sut}[SUT]{System Under Test}
\acro{api}[API]{Application Programming Interfaces}
\acro{3g}[3G]{Third Generation}
\acro{ast}[AST]{Abstract Syntax Tree}
\acro{asts}[ASTs]{Abstract Syntax Trees}
\acro{tci}[TCI]{Test Control Interface}
\acro{tri}[TRI]{Test Runtime Interface}
\acro{mtc}[MTC]{Main Test Component}
\acro{ptc}[PTC]{Parallel Test Component}
\acro{antlr}{Another Tool for Language Recognition}
\acro{trex}[TRex]{TTCN-3 Refactoring and Metrics Tool}
\acro{dot}[DOT]{Drawing of Directed Graphs}
\acro{uml}[UML]{Unified Modeling Language}
\acro{csv}[CSV]{Comma-Separated Values}
\acro{tle}[TLE]{Top Level Element}
\acro{tlecand}[TLEcand]{Top Level Element Candidate}
\end{acronym}