### 2011-04-03 paper.tex

%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage[printonlyused]{acronym}

\usepackage{color}

\newcounter{countNotes}
\newcommand{\annKB}[1]{\color{blue}*\textsuperscript{\thecountNotes}\marginpar{\color{blue} \tiny *\textsuperscript{\thecountNotes}\textbf{Note (KB): \\}
    #1  }\addtocounter{countNotes}{1}\color{black}}
%\renewcommand{\annKB}[1]{} %disable annotation

% \renewcommand{sbw}{} %disable comments

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

\journal{has to be discussed}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Development and Data Collection of a Facebook application}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[focal]{Kathrin Becker\corref{cor1}\fnref{fn1}} 
\ead{mail@kathrin-becker.eu}
\author[focal]{Mayutan Arumaithurai\fnref{fn1}} %
\ead{mayutan.arumaithurai@gmail.com}
\author[focal]{Xiaoming Fu\corref{cor2}\fnref{fn1}} %
\ead{fu@cs.uni-goettingen.de}

\address[focal]{Institute of Computer Science, Computer Networks (NET)
  Research Group, University of G\"ottingen}

\cortext[cor1]{Corresponding author}
\cortext[cor2]{Principal corresponding author}
\fntext[fn1]{This document is a collaborative effort.}
%\fntext[fn2]{Another author footnote, but a little more longer.}
%\fntext[fn3]{Yet another author footnote. Indeed, you can have
%xany number of author footnotes.}
\tnotetext[tn1]{This is a specimen title.}

\begin{abstract}
%% Text of abstract, 150 - 350 words
% The outcome was students' perceptions of preparedness.

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
Facebook, Social Networking, data mining, social graph
\end{keyword}

\end{frontmatter}

%%

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% main text
%% Start line numbering here if you want
%%
%o \linenumbers

\section{Introduction}
\label{sec:introduction}

\section{Methodology}
\label{sec:method}

% Describes:
% \begin{itemize}
% \item Probands (studied objects): Type, number, characteristics, age,
% gender by arithmetic means and distribution
% \item materials: surveys (exemplary demonstration!) incl instructions, software used, given media (differences)
% \item Proceeding: Splitting of the group, quality of questions
% \item Evaluationmethods: methods used for evaluation (statistical tests)
% \end{itemize}

\section{Results}
\label{sec:results}

\begin{center}
\begin{table}
\label{tab:coruser}
\begin{tabular}{l | c c c}
a & a & a & a \\

\hline
b & b & b & b
\end{tabular}
\caption{dummy}
\end{table}
\end{center}

\section{Discussion}
\label{sec:discussion}

\section{Outlook}
\label{sec:outlook}

\section{Conclusion}
\label{sec:conclusion}

\include{acronym} % load acronyms
%% \section{}
%% \label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}          ==>>  [#]
%%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
%%   \citet{key}         ==>>  Author [#]

%% References with bibTeX database:

\bibliographystyle{elsarticle-harv}
\bibliography{literature.bib}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model1-num-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file `elsarticle-template-1-num.tex'.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: